---
title: "HW 1 part 1"
date: "13 02 2023"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(reshape2)
library(kableExtra)
library(Matrix)
library(MASS)
library(mice)
library(Hmisc)
library(corrplot)
library(performance)
library(naniar)
library(psych)
library(GGally)
library(campfin)

knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

# Critical Thinking Group 1 (Ben Inbar, Cliff Lee, Daria Dubovskaia, David Simbandumwe, Jeff Parks, Nick Oliver)

## 1. DATA EXPLORATION


<b><i>
Notes / Questions

- added boxplot and density plot 
- do we want to use a corr plot with the numbers
  - there are a couple of variable with high correlations that we might want to explore
  - Batting_hr - Pitching_hr = .97
  - Batting_hr - Batting_so = .73
  - what is the cutoff in correlation that will require adjustment
-

</i></b>


```{r initialization, include=FALSE}

trainDf <- read.csv("https://raw.githubusercontent.com/cliftonleesps/data621_group1/main/hw1/moneyball-training-data.csv")
evalDf <- read.csv("https://raw.githubusercontent.com/cliftonleesps/data621_group1/main/hw1/moneyball-evaluation-data.csv")

```

The baseball training dataset contains 2,276 observations detailing various teams' performances per year from 1871 to 2006. Due to the relatively long period, we expect to see outliers and missing data as the league modified official game rules; these rule changes undoubtedly caused teams and players to change their tactics in response. Additionally, the number of single base hits is noticeably missing from the columns. However, we will derive this value as the number other types of hits (double, triples, home runs) can be subtracted from total hits. Lastly, other columns representing game number (out of 162), inning number (1-9), and matching opponent columns would have been vastly useful for predictions. One last noticeable omission from the original dataset is the of the number of single base hits. However, this value can possibly be calculated as a difference between other types of hits (double, triples, home runs) and total hits.

#### Summary Statistics

The table below shows us some valuable descriptive statistics for the training data. We can see that many of the variables have a minimum of 0 but not all. The means and medians of each variable are all relatively close in value for each individual variable. This tells us that most data is free from extreme outliers as they tend to skew the mean relative to the median.

```{r summary_chart, echo=FALSE}
stats <- trainDf %>%
  dplyr::select(-c(INDEX)) %>%
  describe()
displayDf <- as.data.frame(stats) %>%
  dplyr::mutate(missing = 2276 - n) %>%
  dplyr::select(-c(vars, n, trimmed, mad, range))
kable(displayDf, digits = 2, format = "html", row.names = T) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = F,
    position = "center"
  )
```
One interesting piece of information is the min/max of the `TARGET_WINS` variable. The minimum is 0 meaning there are teams that did not win a single game. The maximum is 146 which indicates no team in the training dataset had a perfect season, as we know from the data a season consists of 162 games.

Also of note is the number of missing values from certain variables. Most notably the `TEAM_BATTING_HBP` (batters hit by pitch variable). With 91% of the data missing we will remove this variable from our dataset because there simply is not enough information to impute a sensible value. The missing data for these two columns may be due to a change official rules or tactics before the modern era of baseball.

The means and medians of each variable are all relatively close in value for each individual variable. This tells us that most data is free from extreme outliers as they tend to skew the mean relative to the median.



#### Distribution and Box Plots

Next, we'll visually check for normal distributions and box plots in both the dependent and independent variables. The grid plot below shows normalcy in most features except for extremely right skewed features such as hits allowed (PITCH_H) and errors (FIELDING_E).  Box plots for these further show a high number of outliers exist outside of the interquartile ranges so their effects should be carefully considered.

```{r density_boxplot, echo=FALSE}

m_df <- melt(trainDf)
m_df %>% ggplot(aes(x= value)) + 
    geom_density(fill='gray') + facet_wrap(~variable, scales = 'free') 


m_df %>% ggplot(aes(x = value)) +
  geom_boxplot(outlier.color = 'red', outlier.shape = 1) +
  facet_wrap(vars(variable),scales = "free", ncol = 4)

```

Doing some preliminary analysis of the data we can compare the team pitching hits (`TEAM_PITCHING_H`) against the target wins (`TARGET_WINS`) to validate our assumption that more hits allowed is negatively correlated with the number of wins. While there does appear to be a negative correlation it is not obviously linear. There is clearly a lot of clustering below 5,000 allowed hits but the number of wins varies significantly. There are also some obvious outliers here with one team allowing nearly 30,000 hits.

```{r scatter_hits_wins, echo=FALSE}
trainDf %>%
  ggplot(aes(x = TEAM_PITCHING_H, y = TARGET_WINS)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Allowed Hits vs. Target Wins", x = "Allowed Hits", y = "Target Wins")
```

Looking at another relationship between strike outs and wins we can see a the strike out data has fewer outliers but not as strong of a negative correlation as one would expect with target wins. 
```{r, so_wins, echo=FALSE}
trainDf %>%
  ggplot(aes(x = TEAM_BATTING_SO, y = TARGET_WINS)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Strike outs vs. Target Wins", x = "Strike outs", y = "Target Wins")
```


#### Correlation Matrix

Plotting the correlations between TARGET_WINS and the variables (excluding INDEX and TEAM_BATTING_HBP) we can see that very few variables are strongly correlated with the target variable.

```{r correlation_matrix, echo=FALSE}
rcore <- rcorr(as.matrix(trainDf %>% dplyr::select(where(is.numeric) & -INDEX & -TEAM_BATTING_HBP)))
coeff <- rcore$r
corrplot(coeff, tl.cex = .7 , method = 'circle')
```


Comparing offensive (any column starting with BATTING or BASERUN) to defensive stats unexpectedly shows some correlation, pointing to potential problems. Qualitatively, the matrix implies some teams or players are exceptional both at hitting (offensive) and fielding (defensive). Furthermore, a typical team's number of batted home runs and allowed home runs has a correlation of nearly 1.0! This is obviously ridiculous; if a team hits four home runs, is it likely it will also allow its opponent to hit four also? As we know, many baseball games have low even scores (1-2, 3-2) but there many one sided wins also. Home runs hit (BATTING_HR) and home runs allowed (PITCHING_HR) are immediately suspect and are good candidates for outright exclusion. Alternatively, the correlation between a team's hits (BATTING_H) and hits allowed (PITCHING_H) is around 0.3 which is seems reasonable.

There are some other strong correlations that are less obvious such as Errors (`TEAM_FIELDING_E`) being strongly negatively correlated with walks by batters (`TEAM_BATTING_BB`), strike outs (`TEAM_BATTING_SO`).  All combined together, teams that get a lot of hits do not generally make fielding errors.

Digging a little deeper we can see there is a Pearson correlation coefficient of `r cor(trainDf$TEAM_FIELDING_E, trainDf$TEAM_BATTING_BB)` for errors and walks by batters which indicates a strong negative correlation between the two variables. Looking at errors compared with team pitching hits allowed we see a correlation of `r cor(trainDf$TEAM_FIELDING_E, trainDf$TEAM_PITCHING_H)` which indicates a strong positive correlation. 

<b><i>
Notes / Questions

- Given the number of games in a year I don't believe that 0 is a valid measurement for most of the attributes
- use MICE to fill missing values vs Median or Mean


</i></b>





<b><i>
Notes / Questions

- Review the distributions plots for each attribute focus on non normally distributed attributes
- Box plots show outliers discuss valid data based on https://www.baseball-reference.com/leagues/majors/bat.shtml

</i></b>




Lastly lets take a closer look at the missing data. We've already determined that the batter hit by pitch (`TEAM_BATTING_HBP`) variable is missing 91% of its data but what of the other variables.

Using the plot below we can visualize the missingness of the remaining variables. There are 5 variables that contain varying degrees of missing data. We will use the information to fill in the missing values in our data preparation step.

`TEAM_BASERUN_CS` appears to be missing the second most amount of values but at only 772 missing values out of 2276 this is much less of a concern than the HBP variable we identified earlier. The remaining variables that are missing data have less than 25% of their data missing so should be safe to impute.

```{r missing_data, echo=FALSE}
gg_miss_var(trainDf)
```
