---
title: 'Final: Texas Higher Education Opportunity Project (THEOP)'
subtitle: 'Critical Thinking Group 1'
author: 'Ben Inbar, Cliff Lee, Daria Dubovskaia, David Simbandumwe, Jeff Parks'
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: united
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: sentence
---


```{r setup, include=FALSE}
# chunks
knitr::opts_chunk$set(echo=FALSE, eval=TRUE, include=TRUE, 
message=FALSE, warning=FALSE, fig.height=5, fig.align='center')

# libraries
library(tidyverse)
library(kableExtra)
library(MASS) # glm.nb()
library(mice)
library(pscl) # zeroinfl()
library(skimr)
library(sjPlot)
library(mpath)
library(yardstick)
library(labelled)
library(haven)
library(corrplot)
library(Hmisc)
library(janitor)

# ggplot
theme_set(theme_light())


# random seed
set.seed(42)
```





```{r common functions}

nice_table <- function(df){
  table <- df %>% kable %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                html_font = 'monospace',
                full_width = FALSE)
  return(table)
}

model_diag <- function(model){
  model_sum <- summary(model)
  aic <- AIC(model)
  ar2 <- model_sum$adj.r.squared
  disp <- sum(resid(model,'pearson')^2)/model$df.residual
  loglik <- logLik(model)
  
  vec <- c(ifelse(is.null(aic), NA, aic),
           ifelse(is.null(ar2), NA, ar2),
           ifelse(is.null(disp), NA, disp),
           ifelse(is.null(loglik), NA, loglik))
  
  names(vec) <- c('AIC','Adj R2','Dispersion','Log-Lik')
  return(vec)
}



factor_haven <- function(df, col_lst) {
  for (c in col_lst) {
    df[c] <- as_factor(zap_missing(df[c]))
  }
  return(df)
}


```





# Load Data

```{r}

# change to your local data dir outside the repo
# local_data_dir <- '../../data/theop'
# 
# 
# # load application and transactions data frames
# load(paste0(local_data_dir, '/data_model/df_applications.RData'))
# load(paste0(local_data_dir, '/data_model/df_transcripts.RData'))

load("C:\\Users\\bpinb\\OneDrive\\Desktop\\Coursework\\DATA 621 Business Analytics and Data Mining\\Final project\\Data\\theop\\theop\\data_model\\df_applications.RData")
load("C:\\Users\\bpinb\\OneDrive\\Desktop\\Coursework\\DATA 621 Business Analytics and Data Mining\\Final project\\Data\\theop\\theop\\data_model\\df_transcripts.RData")

```





```{r}

# clean application data and remove labels
app_df  <- df_applications
col_lst <- c("termdes","male","ethnic","citizenship","restype","satR","actR","testscoreR","decileR","quartile","major_field",
             "hsprivate","hstypeR","hsinstate","hseconstatus","hslos","hscentury","admit","admit_prov","enroll","gradyear",
             "studentid_uniq","univ","termapp","sat_not_recenteredR","admit_ut_summer")
app_df <- factor_haven(app_df,col_lst)


# clean transcripts and remove labels
transcript_df <- df_transcripts
col_lst <- c("term","semgpa","hrearn","term_major_dept","term_major_field")
transcript_df <- factor_haven(transcript_df,col_lst)
rm(df_transcripts, df_applications)
```






# Data Exploration


```{r}

app_df %>% 
  #dplyr::select(!c(source, target)) %>%
  skim() %>%
  dplyr::select(skim_variable, complete_rate, n_missing, 
                numeric.p0, numeric.p100) %>%
  rename(variable=skim_variable, min=numeric.p0, max=numeric.p100) %>%
  mutate(complete_rate=round(complete_rate,2), 
         min=round(min,2), max=round(max,2)) %>%
  arrange(variable) %>%
  nice_table()


transcript_df %>% 
  #dplyr::select(!c(source, target)) %>%
  skim() %>%
  dplyr::select(skim_variable, complete_rate, n_missing, 
                numeric.p0, numeric.p100) %>%
  rename(variable=skim_variable, min=numeric.p0, max=numeric.p100) %>%
  mutate(complete_rate=round(complete_rate,2), 
         min=round(min,2), max=round(max,2)) %>%
  arrange(variable) %>%
  nice_table()

```




```{r distrib}

transcript_df %>% dplyr::select(where(is.numeric)) %>%
  #dplyr::select(!target) %>%
  pivot_longer(everything(),names_to = c('variables'),values_to = c('values')) %>% 
  ggplot() +
  geom_histogram(aes(x=values, y = ..density..), alpha=0.5, colour='black', linewidth=0.2) +
  geom_density(aes(x=values), color='purple') +
  facet_wrap(vars(variables), scales="free")


```




```{r}

rcore <- rcorr(as.matrix(transcript_df %>% dplyr::select(where(is.numeric))))
coeff <- rcore$r
corrplot(coeff, tl.cex = .5, tl.col="black", method = 'color', addCoef.col = "black",
         type="upper", order="hclust", number.cex=0.7, diag=FALSE)

```



# Transform Data

```{r}
t_df <- transcript_df

t_df$term_num <- as.numeric(t_df$term)
#t_df <- t_df %>% mutate(id = id + 1)
t_df$term_num = t_df$term_num + 1
t_df$term_num[t_df$term_num == 2] <- 1
t_df$term_num[t_df$term_num == 6] <- 2
t_df$term_num[t_df$term_num == 7] <- 6
t_df$term_num = t_df$term_num / 10
t_df$term_num = t_df$year + t_df$term_num


avg_df <- t_df %>%
  group_by(studentid) %>%
  summarise(avg_gpa = mean(cgpa)
 )

first_df <- t_df %>%
  group_by(studentid) %>%
  slice(which.min(term_num))
  
last_df <- t_df %>%
  group_by(studentid) %>%
  slice(which.max(term_num))

merge_df <- merge(avg_df, first_df, by='studentid')
merge_df <- merge(merge_df, last_df, by='studentid', suffixes = c(".first",".last"))

merge_df <- merge(merge_df, app_df, by='studentid')


```

# Merge dataframes. Check for duplicates in applications and transcripts. For applications there were duplicates for the desired year of admission, so we took the minimum desired year. For transcripts, make sure there are no duplicates for the same semester. We also limited to enrollment at the applied school to track progress only at schools of enrollment.
```{r}
tabyl(app_df$termdes)
#dupes <- get_dupes(app_df, studentid_uniq, enroll)

app_df <- app_df |>
  mutate(termdes_num = case_when(termdes=='Spring' ~ 1
                                 , termdes=='Summer I' ~ 2
                                 , termdes=='Summer II' ~ 3
                                 , termdes=='Fall' ~ 4)) |>
  group_by(studentid_uniq) |>
  filter(yeardes==min(yeardes)) |>
  filter(termdes_num==min(termdes_num)) |>
  filter(enroll=="Yes")
  
  
dupes <- get_dupes(transcript_df, studentid_uniq, year, term)

transcript_df <- transcript_df |>
  distinct()

# Join them together
all <- app_df |>
  left_join(transcript_df, by='studentid_uniq')
  
  

```


```{r}

merge_df %>% 
  #dplyr::select(!c(source, target)) %>%
  skim() %>%
  dplyr::select(skim_variable, complete_rate, n_missing, 
                numeric.p0, numeric.p100) %>%
  rename(variable=skim_variable, min=numeric.p0, max=numeric.p100) %>%
  mutate(complete_rate=round(complete_rate,2), 
         min=round(min,2), max=round(max,2)) %>%
  arrange(variable) %>%
  nice_table()

```




```{r}


rcore <- rcorr(as.matrix(merge_df %>% dplyr::select(!c('year.first','year.last')) %>% dplyr::select(where(is.numeric))))
coeff <- rcore$r
corrplot(coeff, tl.cex = .8, tl.col="black", method = 'color', addCoef.col = "black",
         type="upper", order="hclust", number.cex=0.7, diag=FALSE)


```


